---
author: "Gustavo A. Ballen, Carlos Jaramillo, Fernando C. P. Dagosta & Mario C. C. de Pinna"
title: "A fossil fish assemblage from the middle Miocene of the Cocinetas Basin, northern Colombia"
subtitle: "Supplementary material"
output: 
    pdf_document:
        number_sections: true
        citation_package: natbib
        keep_tex: true
urlcolor: blue
toc: true
biblio-style: apalike
bibliography: PhDThesis.bib
header-includes: 
- \usepackage{longtable}
- \usepackage{lscape}
- \usepackage{inputenc}
- \renewcommand{\thesection}{S\arabic{section}}
- \renewcommand{\thetable}{S\arabic{table}}
- \renewcommand{\thefigure}{S\arabic{figure}}
---

```{r, echo = FALSE}
library(vegan)
library(knitr)
library(pvclust)
library(sf)
opts_chunk$set(tidy.opts = list(width.cutoff = 58), tidy = TRUE)
```

# Faunal similarity: Expanded analysis

```{r}
### faunal similarity analyses for the Marakaipao fish fauna
dataset <- read.delim(file = "neogeneFishOccs.tab", stringsAsFactors = FALSE)

#################### only freswater taxa
dataset <- dataset[which(dataset$Environment == "Freshwater" | dataset$Environment == "Both"), ]

### complete analysis without filtering localities
wholeMatrix <- dataset[, -c(1, 3, 19)]
rownames(wholeMatrix) <- wholeMatrix$Taxon
wholeMatrix <- wholeMatrix[-1]
wholeMatrix <- t(wholeMatrix)

###include only faunas with positive number of occurrences
wholeMatrix <- wholeMatrix[apply(X = wholeMatrix, MARGIN = 1, FUN = sum) > 0, ]

### number of occurrences per fauna
sort(x = apply(X = wholeMatrix, MARGIN = 1, FUN = sum), decreasing = TRUE)

### calculate the distance matrix using Bray-Curtis' method
distMatrixBinarywhole <- pvclust::pvclust(t(wholeMatrix), method.hclust = "average", method.dist = "binary")
```

```{r, fig.asp = 1}
plot(distMatrixBinarywhole, main = "Faunal similarity")
```

# Dataset cleaning

Data from Species Link were downloaded and stored for the whole of Chordata, so that specific names could be sliced with the following scripts:

```{bash, eval = FALSE}
### slicer
#!/usr/bin/env bash

# TAXON and FILE are mandatory arguments for the expression (taxon name) TAXON to be
# serched for by grep in the file FILE.
# the result is written to a new file preserving the original FILE filename and
# appending a .out string so it can be differentiated
TAXON=$1
FILE=$2
OUTPUT=`(basename $FILE)`.out
head -n 1 $FILE > $OUTPUT
grep $TAXON $FILE >> $OUTPUT
NLINES=`wc -l *out | awk '{print $1}'`
echo "Finished writing $OUTPUT with $NLINES lines"
```

```{bash, eval = FALSE}
### speciesLinkPicker.sh
# this script uses slicer, a small program in bash that I wrote for that specific purpose.
# SPLINKFILE is the path to the specieslink file directory
SPLINKFILE=speciesLink_all_49771_20190521013813.txt

# Callichthyidae
slicer Callichthyidae $SPLINKFILE
cp *out callichthyidae/
rm *out

# Lepidosiren
slicer Lepidosiren $SPLINKFILE
cp *out lepidosiren/
rm *out

# Phractocephalus
slicer Phractocephalus $SPLINKFILE
cp *out phractocephalus/
rm *out

# Serrasalmidae
slicer Serrasalmidae $SPLINKFILE
cp *out serrasalmidae/
rm *out
```

## Specific data cleaning

### Callichthyidae

```{R, eval = FALSE}
gbifData <- read.delim("occurrence.txt", header = TRUE, stringsAsFactors =  FALSE)
splinkData <- read.delim("speciesLink_all_49771_20190521013813.txt.out", header = TRUE, stringsAsFactors =  FALSE)

### clean the SpeciesLink dataset
# remove data from specieslinkData with notes containing "bloqueado"
splinkData <- splinkData[-grep(x = splinkData$notes, pattern = "bloqueado"), ]
# the latter case was messing with the type of the coordinate columns, coerce them to numeric
splinkData$longitude <- as.numeric(splinkData$longitude)
splinkData$latitude <- as.numeric(splinkData$latitude)
# coordinate automatic cleaning
# examine the boxplot
boxplot(splinkData$latitude)
boxplot(splinkData$longitude)
# filter out all coordinates east of (South America+Panama)'s
# easternmost point: -34.7930, João Pessoa, Paraíba, and 
splinkData <- splinkData[-which(splinkData$longitude > -34.7930), ]

### Clean the GBIF dataset
# remove cases that are explicitly showing geospatial data
gbifData <- gbifData[-which(gbifData$hasGeospatialIssues == "true"),  ]
# remove all coordinates out of (South America + Panama)'s bounds
gbifData <- gbifData[-which(gbifData$decimalLongitude > -34.7930), ]
gbifData <- gbifData[-which(gbifData$decimalLongitude < -83.0521), ]
gbifData <- gbifData[-which(gbifData$decimalLatitude > 12.4583), ]
# remove an erroneous record of Corydoras from Guajira along with caribbean marine occurrences
gbifData <- gbifData[-which(gbifData$decimalLatitude > 12.20), ]
# remove erroneous records from the Pacific
gbifData <- gbifData[-which(gbifData$decimalLatitude == 3.997866), ]
# remove erroneous records from the Caribbean
gbifData <- gbifData[-which(gbifData$decimalLatitude == 11.616667), ]
# remove erroneous records from the Caribbean
gbifData <- gbifData[-which(gbifData$decimalLatitude == 11), ]
# remove erroneous records from the Atlantic
gbifData <- gbifData[-which(gbifData$decimalLatitude == -16.851667), ]
# remove erroneous records from the Atlantic
gbifData <- gbifData[-which(gbifData$decimalLatitude == -24.134167), ]
# remove erroneous records from the Atlantic
gbifData <- gbifData[-which(gbifData$decimalLatitude == -23.983333), ]
# remove erroneous records from the Atlantic
gbifData <- gbifData[-which(gbifData$decimalLatitude == -24.134167), ]
# remove erroneous records from the Atlantic
gbifData <- gbifData[-which(gbifData$decimalLatitude == -29.943333), ]
# remove fossil occurrences
gbifData <- gbifData[-which(gbifData$basisOfRecord == "FOSSIL_SPECIMEN"),  ]

### Clean the specieslink dataset
# remove an erroneous record of Corydoras from Guajira along with caribbean marine occurrences
splinkData <- splinkData[-which(splinkData$latitude > 12.20), ]
# remove erroneous records from the Atlantic
splinkData <- splinkData[-which(splinkData$latitude == -24.134167), ]
# remove erroneous records from the Atlantic
splinkData <- splinkData[-which(splinkData$latitude == -29.943333), ]

### Taxonomic cleaning
# remove data w/o genus
gbifData <- gbifData[-which(gbifData$genus == ""), ]
splinkData <- splinkData[-which(splinkData$genus == ""), ]
splinkData <- splinkData[-which(splinkData$genus == "Callichthyidae"), ]

### subsetting
# construct the dataframe concatenating column contents and creating on the fly a column with information on the source of the record
callichCoords <- data.frame(genus = c(gbifData$genus, splinkData$genus),
                            species = c(gbifData$species, splinkData$scientificname),
                            latitude = c(gbifData$decimalLatitude, splinkData$latitude),
                            longitude = c(gbifData$decimalLongitude, splinkData$longitude),
                            source = c(rep("gbif", times = nrow(gbifData)),
                                       rep("splink", times = nrow(splinkData))), 
                            stringsAsFactors = FALSE)
# remove missing coordinates. 
# identical(is.na(callichCoords$latitude),
#  is.na(callichCoords$longitude)) evaluates to TRUE
# so we can just pick any of the coordinate components
callichCoords <- callichCoords[!is.na(callichCoords$longitude), ]
#write the clean dataset
write.table(x = callichCoords, file = "callichtyidae.tab", sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
```

### \textit{Lepidosiren}

```{R, eval = FALSE}
gbifData <- read.delim("occurrence.txt", header = TRUE, stringsAsFactors =  FALSE)
splinkData <- read.delim("speciesLink_all_49771_20190521013813.txt.out", header = TRUE, stringsAsFactors =  FALSE)

### clean the SpeciesLink dataset
# coordinate automatic cleaning
# examine the boxplot
boxplot(splinkData$latitude)
boxplot(splinkData$longitude)
# filter out all coordinates east of (South America+Panama)'s
# easternmost point: -34.7930, João Pessoa, Paraíba, and 
splinkData <- splinkData[-which(splinkData$longitude > -34.7930), ]

### Clean the GBIF dataset
# remove cases that are explicitly showing geospatial data
gbifData <- gbifData[-which(gbifData$hasGeospatialIssues == "true"),  ]
# remove fossil occurrences
gbifData <- gbifData[-which(gbifData$basisOfRecord == "FOSSIL_SPECIMEN"),  ]

### Taxonomic cleaning

### subsetting
# construct the dataframe concatenating column contents and creating on the fly a column with information on the source of the record
lepidoCoords <- data.frame(genus = c(gbifData$genus, splinkData$genus),
                            species = c(gbifData$species, splinkData$scientificname),
                            latitude = c(gbifData$decimalLatitude, splinkData$latitude),
                            longitude = c(gbifData$decimalLongitude, splinkData$longitude),
                            source = c(rep("gbif", times = nrow(gbifData)),
                                       rep("splink", times = nrow(splinkData))), 
                            stringsAsFactors = FALSE)
# remove missing coordinates. 
# identical(is.na(lepidoCoords$latitude),
#  is.na(lepidoCoords$longitude)) evaluates to TRUE
# so we can just pick any of the coordinate components
lepidoCoords <- lepidoCoords[!is.na(lepidoCoords$longitude), ]
#write the clean dataset
write.table(x = lepidoCoords, file = "lepidosiren.tab", sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
```

### \textit{Phractocephalus}

```{R, eval = FALSE}
gbifData <- read.delim("occurrence.txt", header = TRUE, stringsAsFactors =  FALSE)
splinkData <- read.delim("speciesLink_all_49771_20190521013813.txt.out", header = TRUE, stringsAsFactors =  FALSE)

### clean the SpeciesLink dataset
# examine the boxplot
boxplot(splinkData$latitude)
boxplot(splinkData$longitude)
# filter out all coordinates east of (South America+Panama)'s
# easternmost point: -34.7930, João Pessoa, Paraíba, and 
splinkData <- splinkData[-which(splinkData$longitude > -34.7930), ]

### Clean the GBIF dataset
# remove cases that are explicitly showing geospatial data
gbifData <- gbifData[-which(gbifData$hasGeospatialIssues == "true"),  ]
# remove all coordinates out of (South America + Panama)'s bounds
gbifData <- gbifData[-which(gbifData$decimalLongitude < -83.0521), ]
gbifData <- gbifData[-which(gbifData$decimalLatitude > 12.4583), ]
# remove fossil occurrences
gbifData <- gbifData[-which(gbifData$basisOfRecord == "FOSSIL_SPECIMEN"), ]
# two doubtful records removed, one b/c of geographic mismatch
# the other because of geographic uncertainty placing the record out of the known distribution
gbifData <- gbifData[-which(gbifData$occurrenceID == "086E2F34-26EC-4270-9CC8-771FFCCA737B"), ]
gbifData <- gbifData[-which(gbifData$occurrenceID == "BR:UEL:MZUEL-Peixes:18116"), ]
# the last case was duplicated in the specieslink dataset
splinkData <- splinkData[-which(splinkData$catalognumber == "17189"),  ]


### Taxonomic cleaning
# remove an erroneous record of Paracanthopoma
splinkData <- splinkData[-which(splinkData$genus == "Paracanthopoma"),  ] 
### subsetting
# construct the dataframe concatenating column contents and creating on the fly a column with information on the source of the record
phractoCoords <- data.frame(genus = c(gbifData$genus, splinkData$genus),
                            species = c(gbifData$species, splinkData$scientificname),
                            latitude = c(gbifData$decimalLatitude, splinkData$latitude),
                            longitude = c(gbifData$decimalLongitude, splinkData$longitude),
                            source = c(rep("gbif", times = nrow(gbifData)),
                                       rep("splink", times = nrow(splinkData))), 
                            stringsAsFactors = FALSE)
# remove missing coordinates. 
# identical(is.na(phractoCoords$latitude),
#  is.na(phractoCoords$longitude)) evaluates to TRUE
# so we can just pick any of the coordinate components
phractoCoords <- phractoCoords[!is.na(phractoCoords$longitude), ]
#write the clean dataset
write.table(x = phractoCoords, file = "phractocephalus.tab", sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
```

### Serrasalmidae

```{R, eval = FALSE}
gbifData <- read.delim("occurrence.txt", header = TRUE, stringsAsFactors =  FALSE)
splinkData <- read.delim("speciesLink_all_49771_20190521013813.txt.out", header = TRUE, stringsAsFactors =  FALSE)

### clean the SpeciesLink dataset
# coordinate automatic cleaning
# examine the boxplot
boxplot(splinkData$latitude)
boxplot(splinkData$longitude)
# filter out all coordinates east of (South America+Panama)'s
# easternmost point: -34.7930, João Pessoa, Paraíba, and 
splinkData <- splinkData[-which(splinkData$longitude > -34.7930), ]

### Clean the GBIF dataset
# remove cases that are explicitly showing geospatial data
gbifData <- gbifData[-which(gbifData$hasGeospatialIssues == "true"),  ]
# remove all coordinates out of (South America + Panama)'s bounds
gbifData <- gbifData[-which(gbifData$decimalLongitude > -34.7930), ]
gbifData <- gbifData[-which(gbifData$decimalLongitude < -83.0521), ]
gbifData <- gbifData[-which(gbifData$decimalLatitude > 12.4583), ]
# remove serrasalmid occurrences in the ocean and west of the Andes in Peru where they are certainly erroneous
# these occurrences are said to be from Leticia, Colombia.
gbifData <- gbifData[-which(gbifData$decimalLatitude == 3.997866), ]
# remove an occurrence from "Pibas" (most likely Pebas) that mapped west of the Andes
gbifData <- gbifData[-which(gbifData$decimalLatitude == -12.454263), ]
# remove an occurrence from west of the Andes of Piaractus from Sucre, introduced to the Magdalena-Cauca
gbifData <- gbifData[-which(gbifData$decimalLatitude == 9.166804), ]
# remove fossil occurrences
gbifData <- gbifData[-which(gbifData$basisOfRecord == "FOSSIL_SPECIMEN"), ]

### Taxonomic cleaning
# remove occurrences uncertain to family level
splinkData <- splinkData[-which(splinkData$genus == ""),  ]
gbifData <- gbifData[-which(gbifData$genus == ""),  ]
### subsetting
# construct the dataframe concatenating column contents and creating on the fly a column with information on the source of the record
serraCoords <- data.frame(genus = c(gbifData$genus, splinkData$genus),
                            species = c(gbifData$species, splinkData$scientificname),
                            latitude = c(gbifData$decimalLatitude, splinkData$latitude),
                            longitude = c(gbifData$decimalLongitude, splinkData$longitude),
                            source = c(rep("gbif", times = nrow(gbifData)),
                                       rep("splink", times = nrow(splinkData))), 
                            stringsAsFactors = FALSE)
# remove missing coordinates. 
# identical(is.na(serraCoords$latitude),
#  is.na(serraCoords$longitude)) evaluates to TRUE
# so we can just pick any of the coordinate components
serraCoords <- serraCoords[!is.na(serraCoords$longitude), ]
#write the clean dataset
write.table(x = serraCoords, file = "serrasalmidae.tab", sep = "\t", row.names = FALSE, fileEncoding = "UTF-8")
```

# Correlates to community similarity

```{R}
### Recalculate the distance matrix for the reduced dataset as presented in the manuscript
### faunal similarity analyses for the Marakaipao fish fauna
dataset <- read.delim(file = "neogeneFishOccs.tab", stringsAsFactors = FALSE)

#################### only freswater taxa
dataset <- dataset[which(dataset$Environment == "Freshwater" | dataset$Environment == "Both"), ]

### Solimões-Pebas does not seem to be a fauna but a collection of different faunas across the Amazon, remove it
dataset <- dataset[, -grep(pattern = "Pebas", x = colnames(dataset))]

comMatrix <- dataset[, -c(1, 3, 18)]
rownames(comMatrix) <- comMatrix$Taxon
comMatrix <- comMatrix[-1]
comMatrix <- t(comMatrix)

### number of occurrences per fauna
sort(x = apply(X = comMatrix, MARGIN = 1, FUN = sum), decreasing = TRUE)

### include only those faunas with at least the number of occurrences
### of Makaraipao
selectFaunas <- names(which(apply(X = comMatrix, MARGIN = 1, FUN = sum) >= apply(X = comMatrix, MARGIN = 1, FUN = sum)["Makaraipao"]))
comMatrix <- comMatrix[selectFaunas, ]

### remove zero-sum species after faunal selection
selectSpp <- names(which(apply(X = comMatrix, MARGIN = 2, FUN = sum) > 0))
comMatrix <- comMatrix[, selectSpp]

### calculate the distance matrix using Bray-Curtis' method
distMatrixBray <- vegan::vegdist(comMatrix, method = "bray", binary = TRUE)
# rename labels in order to replace dots with spaces
attr(distMatrixBray, "Labels") <- gsub(pattern = "\\.", replacement = " ", x = attr(distMatrixBray, "Labels"))
# rename trans-Andean labels in order to place a leading asterisk
attr(distMatrixBray, "Labels") <- gsub(pattern = "Urumaco", replacement = "( T ) Urumaco", x = attr(distMatrixBray, "Labels"), fixed = TRUE)
attr(distMatrixBray, "Labels") <- gsub(pattern = "La Venta", replacement = "( T ) La Venta", x = attr(distMatrixBray, "Labels"), fixed = TRUE)
attr(distMatrixBray, "Labels") <- gsub(pattern = "Makaraipao", replacement = "( T ) Makaraipao", x = attr(distMatrixBray, "Labels"), fixed = TRUE)
```

```{R}
### similarity vs. linear distance
points <- sf::st_read("paleoMiocene.kml")

# solve problems with labels
pairwisePoints <- sf::st_distance(points)
rownames(pairwisePoints) <- points$Name
rownames(pairwisePoints) <- gsub(pattern = " ", replacement = ".", x = rownames(pairwisePoints))
rownames(pairwisePoints) <- gsub(pattern = ".Fm.", replacement = "", x =rownames(pairwisePoints))
rownames(pairwisePoints) <- gsub(pattern = "Acre", replacement = "Rio.Acre", x =rownames(pairwisePoints))
colnames(pairwisePoints) <- points$Name
colnames(pairwisePoints) <- gsub(pattern = " ", replacement = ".", x = colnames(pairwisePoints))
colnames(pairwisePoints) <- gsub(pattern = ".Fm.", replacement = "", x =colnames(pairwisePoints))
pairwisePoints <- as.dist(pairwisePoints)

# sort labels
pairwisePoints <- as.matrix(pairwisePoints)
pairwisePoints <- as.dist(pairwisePoints[order(rownames(pairwisePoints)),order(colnames(pairwisePoints))])

# recalculate the pairwise community distance for easy manipulation
pairwiseComm <- dist(comMatrix, method = "binary")

# sort labels
pairwiseComm <- as.matrix(pairwiseComm)
pairwiseComm <- as.dist(pairwiseComm[order(rownames(pairwiseComm)),order(colnames(pairwiseComm))])

# spearman's rho correlation
corPointsComm <- cor.test(x = pairwisePoints, y = pairwiseComm, method = "spearman", exact = FALSE)

plot(x = pairwisePoints, y = pairwiseComm, xlab = "Geographic distance", ylab = "Community similarity", xlim = c(0, 3500000), pch = 21, bg = "black")

legend(x = "bottomright",
       legend = paste("rho = ",
                      round(corPointsComm$estimate, digits = 3),
                      "; p = ",
                      round(corPointsComm$p.value, digits = 3),
                      sep = ""))
```

```{R}
# similaraity vs. geoAge

# age data
# Urumaco (Urumaco Fm.) ~ 8Ma; 8.0
# Makaraipao (Castilletes Fm.) ~ 15 Ma; 15.0
# La Venta (Honda Gr.) ~ 11 - 13 Ma; 12.0
# Contamana (Pebas-Ipururo Fms.) = middle to late Miocene ~ 14-8 Ma?; 11.0
# Fitzcarrald (No stratigraphic control, Pebas Fm.?) = Laventan ~ 11.8 - 13.8 Ma; 12.8 Ma
# Ituzaingo (Conglomerado Osífero, Ituzaingo Fm.) ~ 6 - 9 Ma; 7.5
# Rio Acre (Solimões Fm.) = Huayquerian ~ 6.8 - 9 Ma; 7.9

# Sorted:
# 07.5 Ituzaingo; 
# 07.9 Rio Acre; 
# 08.0 Urumaco; 
# 11.0 Contamana; 
# 12.0 La Venta; 
# 12.8 Fitzcarrald; 
# 15.0 Makaraipao; 


geoAges <- c(11.0, 12.8, 7.5, 12.0, 15.0, 7.9, 8.0)

names(geoAges) <- c("Contamana", "Fitzcarrald", "Ituzaingo", "La.Venta", "Makaraipao", "Rio.Acre", "Urumaco")

pairwiseAges <- dist(geoAges)

# spearman's rho correlation
corAgesComm <- cor.test(x = pairwiseAges, y = pairwiseComm, method = "spearman", exact = FALSE)

plot(x = pairwiseAges, y = pairwiseComm, xlab = "Geological age distance", ylab = "Community similarity", pch = 21, bg = "black")

legend(x = "bottomright",
       legend = paste("rho = ",
                      round(corAgesComm$estimate, digits = 3),
                      "; p = ",
                      round(corAgesComm$p.value, digits = 3),
                      sep = ""))
```


# Raw occurrence data

\begin{landscape}
\begin{tiny}
\begin{longtable}{llllllllllllllllllp{3cm}}
\caption{Occurrence table. Acronyms of fossil faunas are: Castillo=CS, La Venta=LV, Makaraipao=MK, Castilletes marine=CM, Loyola-Mangan=LM, Rio Acre=RA, Solimões-Pebas=SP, Urumaco=UF, Utuquina=UQ, Pirabas=PI, Cantaure=CA, Ituzaingo=IT, Rio Yuca=RY, Fitzcarrald=FZ, Contamana=CT
}
\label{tab:fossilOccs}\\
Family          & Taxon            & Environment & CS & LV & MK & CM & LM & RA & SP & UF & UQ & PI & CA & IT & RY & FZ & CT & Ref.                                                         \\

\endfirsthead

\endhead

Acregoliathidae & Acregoliath      & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 1           & 0         & \citet{Ballen2019a,Lundberg2010b,Tejada-Lara2015}                 \\
Arapaimidae     & Arapaima         & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Anostomidae     & Leporinus        & Freshwater  & 0        & 1        & 0          & 0                  & 1             & 0        & 1              & 0       & 0        & 0       & 0        & 1         & 0        & 0           & 1         & \citet{Antoine2016,Bogan2012,Lundberg2010b}                       \\
Characidae      & Salminus         & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 1         & 0        & 0           & 0         & \citet{Cione2013}                                                 \\
Serrasalmidae   & Colossoma        & Freshwater  & 1        & 1        & 0          & 0                  & 0             & 1        & 0              & 1       & 0        & 0       & 0        & 1         & 0        & 0           & 0         & \citet{Cione2000,Lundberg2010b}                                   \\
Serrasalmidae   & Megapiranha      & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 1         & 0        & 0           & 0         & \citet{Cione2009}                                                 \\
Serrasalmidae   & Mylossoma        & Freshwater  & 1        & 0        & 1          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Serrasalmidae   & Piaractus        & Freshwater  & 0        & 0        & 1          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Cynodontidae    & Hydrolycus       & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 1         & 0        & 1           & 1         & \citet{Antoine2016,Cione2000,Lundberg2010b,Tejada-Lara2015}       \\
Erythrinidae    & Paleohoplias     & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Erythrinidae    & Hoplias          & Freshwater  & 0        & 1        & 0          & 0                  & 1             & 0        & 1              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Callichthyidae  & Hoplosternum     & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Loricariidae    & Acanthicus       & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Doradidae       & Doraops          & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Doradidae       & Doras            & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Doradidae       & Oxydoras         & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 1        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Doradidae       & Rhinodoras       & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Pimelodidae     & Brachyplatystoma & Freshwater  & 0        & 1        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Aguilera2013d,Lundberg2010b}                               \\
Pimelodidae     & Phractocephalus  & Freshwater  & 0        & 1        & 1          & 0                  & 0             & 1        & 0              & 1       & 0        & 0       & 0        & 1         & 1        & 1           & 1         & \citet{Antoine2016,Azpelicueta2016,Lundberg2010b,Tejada-Lara2015} \\
Pimelodidae     & Platysilurus     & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 1        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Pimelodidae     & Zungaro          & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Ariidae         & Amphiarius       & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Ariidae         & Aspistor         & Both        & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Ariidae         & Bagre            & Marine      & 1        & 0        & 0          & 1                  & 0             & 0        & 0              & 1       & 0        & 0       & 1        & 0         & 0        & 0           & 0         & \citet{Aguilera2013a,Lundberg2010b}                               \\
Ariidae         & Cathorops        & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 1       & 0        & 0         & 0        & 0           & 0         & \citet{Aguilera2013a}                                             \\
Ariidae         & Cantarius        & Marine      & 0        & 0        & 0          & 1                  & 0             & 0        & 0              & 0       & 0        & 0       & 1        & 0         & 0        & 0           & 0         & \citet{Aguilera2013a}                                             \\
Ariidae         & Notarius         & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Ariidae         & Sciades          & Both        & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Ctenosciaena     & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Cynoscion        & Both        & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Equetus          & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Larimus          & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Micropogonias    & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Nebris           & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Ophioscion       & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Pachypops        & Freshwater  & 0        & 0        & 0          & 0                  & 0             & 0        & 1              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Paralonchurus    & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Plagioscion      & Freshwater  & 1        & 0        & 0          & 0                  & 0             & 0        & 1              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Protosciaena     & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sciaenidae      & Xenotolithus     & NA          & 0        & 0        & 0          & 0                  & 0             & 0        & 1              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Serranidae      & Epinephelus      & Marine      & 0        & 0        & 0          & 0                  & 0             & 0        & 0              & 1       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Sphyraenidae    & Sphyraena        & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Scombridae      & Acanthocybium    & Marine      & 1        & 0        & 0          & 0                  & 0             & 0        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 0           & 0         & \citet{Lundberg2010b}                                             \\
Lepidosirenidae & Lepidosiren      & Freshwater  & 0        & 1        & 1          & 0                  & 0             & 1        & 0              & 0       & 0        & 0       & 0        & 0         & 0        & 1           & 1         & \citet{Antoine2016,Lundberg2010b,Tejada-Lara2015}
\end{longtable}
\end{tiny}
\end{landscape}

\clearpage
